---
title: "Model runs using area as cost"
author: "Rachel Rhodes"
date: "12/3/2020"
output: html_document
---
## Time to do our first model runs using area as a cost

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load in packages
library(prioritizr)
library(here)
library(tidyverse)
library(tmap)

```

## First, we must assign planning unit, cost, and conservation features
# Planning Unit
```{r}
## Load in the planning unit raster created in Script 1
pu <- here("rasters", "mz_id.tif")
pu_rast <-  raster(pu)

## print and plot to verify it looks as expected
print(pu_rast)
plot(pu_rast)

```

# Cost
```{r}
## Next, we will want to assign area as cost. Since this did not require that much prep work, we will do the prep work directly in this Rmarkdown

## Pull in the Moz Raster 
area_raster <- raster(here("rasters", "mz_id.tif"))

## Assign uniform cost value of 0.5 to all cells since the area of the plannign units are the same. We assigned a value of 0.5 since we will log transform cost layers in future simulations to be between 0-1 and this 0.5 is a middle value in this range 
values(area_raster) <- 0.5

## Mask the area raster to just the Moz EEZ so all other cells outside of this get assigned a value of NA
area_raster_eez <- mask(area_raster, pu_rast)

## Make a data frame to check values
area_df <- as.data.frame(area_raster_eez)

## Plot to make sure it looks good
plot(area_raster_eez)


## Assign the area_raster as cost which we will call "cost_area"
cost_area <- area_raster_eez

```

# Conservation Features
```{r}
## Create a raster stack from the multilayer habitat raster and the species distirbution tifs
features_stack = stack(here("rasters", "multilayer_feature.tif"), here("rasters", "species", "species_Fis-31568.tif"),  here("rasters", "species", "species_Fis-23898.tif"), here("rasters", "species", "species_Fis-23899.tif"), here("rasters", "species", "species_Fis-23061.tif"), here("rasters", "species", "species_Fis-23064.tif"), here("rasters", "species", "species_Fis-58485.tif"), here("rasters", "species", "species_Fis-29423.tif"), here("rasters", "species", "species_Fis-29388.tif"), here("rasters", "species", "species_Fis-30583.tif"), here("rasters", "species", "species_Fis-23273.tif"), here("rasters", "species", "species_Fis-23274.tif"), here("rasters", "species", "species_Fis-23277.tif"), here("rasters", "species", "species_Fis-8339.tif"), here("rasters", "species", "species_Fis-47352.tif"), here("rasters", "species", "species_Fis-24098.tif"), here("rasters", "species", "species_Fis-163295.tif"), here("rasters", "species", "species_Fis-61508.tif"), here("rasters", "species", "species_Fis-30521.tif"), here("rasters", "species", "species_Fis-32599.tif"), here("rasters", "species", "species_Fis-131821.tif"), here("rasters", "species", "species_Fis-32975.tif"))

## make a data frame of the raster stack to double check everything looks as expected 
con_features_df <- as.data.frame(features_stack)

```

```{r}
## We also need to read in the mpa and oil tifs to use we created in Script 1

## MPAS:
mpa_raster<- raster(here("rasters", "mpa.tif")) 

## Oil
oil_raster <- raster(here("rasters", "oil.tif")) 
```

## Now it's time to start building the problem or simulations in our model

# SCENARIO 1: Baseline - 20% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objctive
prob_area <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Print the problem to verify it is as expected   
#print(prob_area)

## Solve problem
sprob_area <- solve(prob_area)

## Plot the solution to see what it looks like
plot(sprob_area,  main = c("Area- 20% Targets"))

```

# SCENARIO 2: Boundary Penalty = 0.5

```{r}
### Now lets try setting a boundary length of 0.5
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0.5

prob_area_blm <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_boundary_penalties(penalty=0.5) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Print the problem to verify it is as expected   
print(prob_area_blm)

## Solve problem
sprob_area_blm <- solve(prob_area_blm)

## Plot the solution to see what it looks like
plot(sprob_area_blm,  main = c("BLM = 0.5"))

```
# SCENARIO 3: No locked in or locked out constraints
```{r}
## SENSITIVTY ANALSYSIS TEST: Now lets run a sensitivity analysis with nothing locked in or out
### OBJECTIVE: Minimize area while meeting 20% representation targets, without locking in MPAs, locking out oil rigs, and a boundary penalty of 0

## Establish the problem
prob_area_no_mpas <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Print the problem to verify
print(prob_area_no_mpas)

## Solve problem
sprob_area_no_mpas <- solve(prob_area_no_mpas)

## Lot to see what it looks like
plot(sprob_area_no_mpas,  main = c("Sensitivty Analyis: Min Area with no MPAs"))
```

# SCENARIO 4: 10% Targets
```{r}
## SENSITIVTY ANALSYSIS TEST: Now lets run a sensitivity analysis with 10% targets
### OBJECTIVE: Minimize area while meeting 10% representation targets, without locking in MPAs, locking out oil rigs, and a boundary penalty of 0

prob_area_10 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve the problem
sprob_area_10<- solve(prob_area_10)

## Plot the results
plot(sprob_area_10,  main ="10% Targets")

```


# SCENARIO 5: 50% Targets
```{r}
## SENSITIVTY ANALSYSIS TEST: Now lets run a sensitivity analysis with 50% targets
### OBJECTIVE: Minimize area while meeting 50% representation targets, without locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Establish the problem
prob_area_50<- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve the problem
sprob_area_50<- solve(prob_area_50)

## Plot to see results
plot(sprob_area_50,  main ="Area - 50% Target w MPAs")
```


```{r}
# Now lets make a restuls graph comparing the different targets 10, 20 and 50
plotstack <- stack(sprob_area_10, sprob_area, sprob_area_50)

tm_shape(plotstack)+
tm_raster(palette = c("#c6c5c5", "#409a00"), n=2, legend.show = FALSE) +
  tm_layout(title = c("(a)", "(b)", "(c)"))

tmap_save(filename = "outputs/varying_targets.png")

```


