---
title: "Model runs using artisinal fishing pressure as cost"
output: html_document
---

## Now lets to do our next set of runs using artisinal fishing pressure as a cost
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(prioritizr)
library(here)
library(tidyverse)

```

## Again, we must assign planning unit, cost, and conservation features

```{r}
## Make sure you have already run script 6 - this will load the PU and Conservation Features in your local environment

## Now we need to load a new cost layer
## Read in  artisanal fisheries raster from tif file
fishing_raster <- raster(here("rasters", "fishing.tif"))

## Create a data frame to check values
fishing_df <- as.data.frame(fishing_raster)

## Let's scale all values 0-1 by dividing by max value using log-transformed +1 fishing values (+1 so that zero values are transformed to zero not NA)
fishing<- log(fishing_raster+1)/cellStats(log(fishing_raster+1), max)

## Create another data frame to make sure it looks good
fishing_df_scaled <- as.data.frame(fishing)

## Plot to make sure it looks good
plot(fishing)

## we are ready to assign cost
cost_fish <- fishing
```


## Now it's time to start building the problem or simulations for these runs using similar parameters from script 5

# SCENARIO 1: Baseline - 20% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize fishing pressure while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_fish <- problem(cost_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)
  
## Solve problem
sprob_fish<- solve(prob_fish) 

## Plot the solution to see what it looks like
plot(sprob_fish,  main ="Baseline: Fishing Pressure - 20% Target")

```


# SCENARIO 2: TEST BOUNDARIES

# Boundary Penalty = 0.5
```{r}
### Now lets try setting a boundary length of 0.5
### OBJECTIVE: Minimize fishing pressure while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penalty of 0.5

prob_fish_blm <- problem(cost_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_boundary_penalties(penalty=0.5) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Print the problem to verify it is as expected   
print(prob_fish_blm)

## Solve problem
sprob_fish_blm <- solve(prob_blm_fish)

## Plot the solution to see what it looks like
plot(sprob_fish_blm,  main = c("BLM = 0.5"))
```

# Boundary penalty = 1
```{r}
### Now lets try setting a boundary length of 1
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 1

prob_fish_blm_1 <- problem(cost_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_boundary_penalties(penalty=1) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Print the problem to verify it is as expected   
print(prob_fish_blm_1)

## Solve problem
sprob_fish_blm_1 <- solve(prob_fish_blm_1)

## Plot the solution to see what it looks like
plot(sprob_fish_blm_1,  main = c("BLM = 1"))

```

# Boundary penalty = 10
```{r}
### Now lets try setting a boundary length of 10
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 1

prob_fish_blm_10 <- problem(cost_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_boundary_penalties(penalty=1) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Print the problem to verify it is as expected   
print(prob_fish_blm_10)

## Solve problem
sprob_fish_blm_10 <- solve(prob_fish_blm_10)

## Plot the solution to see what it looks like
plot(sprob_fish_blm_10,  main = c("BLM = 10"))

```

# SCENARIO 3: No locked in or locked out constraints
```{r}
## SENSITIVTY ANALSYSIS TEST: Now lets run a sensitivity analysis with nothing locked in or out
### OBJECTIVE: Minimize fishing pressure while meeting 20% representation targets, without locking in MPAs, locking out oil rigs, and a boundary penalty of 0

## Establish the problem
prob_no_mpas <- problem(cost_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

print(prob_no_mpas)

#solve problem
sprob_no_mpas <- solve(prob_no_mpas)

plot(sprob_no_mpas,  main = c("Fishing Pressure - no MPAs"))
```

# SCENARIO 4: 10% Targets
```{r}
## SENSITIVTY ANALSYSIS TEST: Now lets run a sensitivity analysis with 10% targets
### OBJECTIVE: Minimize area while meeting 10% representation targets, without locking in MPAs, locking out oil rigs, and a boundary penalty of 0

## Establish the problem
prob_10<- problem(cost_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve the problem
sprob_10<- solve(prob_10)


## Plot the problem to see results
plot(sprob_10,  main ="Fishing Pressure - 50% Target w MPAs")
```

# SCENARIO 5: 50% Targets
```{r}
## SENSITIVTY ANALSYSIS TEST: Now lets run a sensitivity analysis with 50% targets
### OBJECTIVE: Minimize area while meeting 50% representation targets, without locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Establish the problem
prob_50<- problem(cost_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve the problem
sprob_50<- solve(prob_50)

## Plot to see results
plot(sprob_50,  main ="Fishing Pressure - 50% Target w MPAs")

```

#SCENARIO 6: 20% with Spearnose skate removed (Rostroraja alba) FIS-131821

```{r}

# Conservation Features without spearnose skate
# Create a raster stack from the multilayer habitat raster and the species distirbution tifs
features_stack_nospear = stack(here("rasters", "multilayer_feature.tif"), here("rasters", "species", "species_Fis-31568.tif"),  here("rasters", "species", "species_Fis-23898.tif"), here("rasters", "species", "species_Fis-23899.tif"), here("rasters", "species", "species_Fis-23061.tif"), here("rasters", "species", "species_Fis-23064.tif"), here("rasters", "species", "species_Fis-58485.tif"), here("rasters", "species", "species_Fis-29423.tif"), here("rasters", "species", "species_Fis-29388.tif"), here("rasters", "species", "species_Fis-30583.tif"), here("rasters", "species", "species_Fis-23273.tif"), here("rasters", "species", "species_Fis-23274.tif"), here("rasters", "species", "species_Fis-23277.tif"), here("rasters", "species", "species_Fis-8339.tif"), here("rasters", "species", "species_Fis-47352.tif"), here("rasters", "species", "species_Fis-24098.tif"), here("rasters", "species", "species_Fis-163295.tif"), here("rasters", "species", "species_Fis-61508.tif"), here("rasters", "species", "species_Fis-30521.tif"), here("rasters", "species", "species_Fis-32599.tif"), here("rasters", "species", "species_Fis-32975.tif"))

## make a data frame of the raster stack to double check everything looks as expected 
con_features_df_nospear <- as.data.frame(features_stack_nospear)

# Establish new problem without spearnose skate
prob_fish_nospear<- problem(cost_fish, features = features_stack_nospear) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(mpa_raster) %>%
  add_locked_out_constraints(oil_raster) %>%
  add_gurobi_solver(gap = 0.1, time_limit = 1200)

## Solve the problem
sprob_fish_nospear <- solve(prob_fish_nospear)

## Plot to see results
plot(sprob_fish_nospear,  main ="Fishing Pressure - 20% Target, without spearnose skate")

```


## Now lets plot results next to each other to compare
```{r}
# Now lets make a results graph comparing the different targets 10, 20 and 50
plotstack <- stack(sprob_10, sprob, sprob_50)

tm_shape(plotstack)+
tm_raster(palette = c("#c6c5c5", "#409a00"), n=2, legend.show = FALSE) +
  tm_layout(title = c("(a)", "(b)", "(c)"))

tmap_save(filename = "outputs/fishing_varying_targets.png")

```

## Now let's plot boundary penalties, 0.5, 1 and 10 results next to each other to compare
```{r}
plotstack_fish_blm <- stack(sprob_fish_blm, sprob_fish_blm_1, sprob_fish_blm_10)

tm_shape(plotstack_fish_blm)+
tm_raster(palette = c("#c6c5c5", "#409a00"), n=2, legend.show = FALSE) +
  tm_layout(title = c("(a)", "(b)", "(c)"))

tmap_save(filename = "outputs/fish_blms.png")

```
